DATA 621 - Business Analytics and Data Mining

### 2020-11-02 HW4/HW4_Draft.Rmd
Finish first draft

```{r loadData, include=FALSE}
# Load necessary libraries
library(corrplot)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(data.table)
# Set master seed
set.seed(71554)

@@

```{r model1DummyFreq}
trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
trnDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_AMT,
                        fullRank = TRUE, data = trnX)
trnXFD <- predict(trnDumFreq, newdata = trnX)
```

#### Training
At this point the code will be set up to preprocess the data and train the
model. The two pre-processing steps which will be done will be to check for
near-zero variance between the predictors and then use kNN imputations for
missing values. The actual code may be found in the CODE APPENDIX.

```{r model1RunFreq, cache = 2}
trC <- trainControl(method = 'none',            # Using StepAIC
                    classProbs = TRUE,          # Classification
                    summaryFunction = prSummary # Precision/Recall)
)
set.seed(487)
m1FreqFit <- train(x = trnXFD, y = trnFreq, preProcess = c('nzv', 'knnImpute'),
                   trControl = trC, method = 'glmStepAIC',
                   family = binomial(link = 'logit'), direction = 'both',
                   trace = 0)
kable(summary(m1FreqFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 1 Frequency Regression Output")
```

@@

```{r model1DummySev}
trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
trnDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_FLAG,
                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
```

```{r runModel1Sev}
trC <- trainControl(method = 'none')
set.seed(878)
m1SevFit <- train(x = trnXSD, y = trnSev, preProcess = c('nzv', 'knnImpute'),
                   trControl = trC, method = 'glmStepAIC',
                   family = gaussian, direction = 'both',
                   trace = 0)
kable(summary(m1SevFit)$coefficients, digits = 3L,
      caption = "Model 1 Severity Regression Output")
```

@@

# SELECT MODELS
## Model Selection Criteria
For that *classification* algorithm, we will look at accuracy, AUC, and F1, and
select the model which performs best two out of three in those metrics. For the
*regression* algorithm we will look at RMSE, AIC. and MAE (since the values must
be positive) and select the model which performs best two out of three in those
metrics.

### Frequency Models

```{r model1Test}

```

@@

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r zeroVal}
```
```{r statsN}
```
```{r graphsN}
```
```{r graphsF}
```
```{r corrgram}
```
```{r dataClean}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r model1DummyFreq}
```
```{r model1RunFreq}
```
```{r model1DummySev}
```
```{r model1RunSev}
```

>>>>

```{r loadData, include=FALSE}
# Load necessary libraries
library(corrplot)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(data.table)
# Set master seed
set.seed(71554)

@@

```{r model1DummyFreq}
# Cast target as factor for caret purposes
trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
# Set up dummy variables
trnDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_AMT,
                        fullRank = TRUE, data = trnX)
# Create matrix of dummy values
trnXFD <- predict(trnDumFreq, newdata = trnX)
```

#### Training
At this point the code will be set up to preprocess the data and train the
model. The two pre-processing steps which will be done will be to check for
near-zero variance between the predictors and then use kNN imputations for
missing values. Even though the `caret` package allows the preprocessing to
occur in the call to `train`, it will be seperated here for clarity. The actual
code may be found in the CODE APPENDIX.

```{r model1RunFreq, cache = 2}
# preProcess with near-zero value and kNN imputation
trnXFDpp <- preProcess(trnXFD, method = c('nzv', 'knnImpute'), k = 5)

# Create processed predictors
trnXFDI <- predict(trnXFDpp, newdata = trnXFD)
trC <- trainControl(method = 'none',            # Using StepAIC
                    classProbs = TRUE,          # Classification
                    summaryFunction = prSummary # Precision/Recall
)
set.seed(487)
# Train Model
m1FreqFit <- train(x = trnXFDI, y = trnFreq, trControl = trC,
                   method = 'glmStepAIC', family = binomial(link = 'logit'),
                   direction = 'both', trace = 0)
# Print results
kable(summary(m1FreqFit$finalModel)$coefficients, digits = 3L,
      caption = "Model 1 Frequency Regression Output")
```

@@

```{r model1DummySev}
# Extract those observations with values > 0
trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
# Set up dummy variables
trnDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_FLAG,
                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
# Create matrix of dummy values
trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
# preProcess with near-zero value and kNN imputation
trnXSDpp <- preProcess(trnXSD, method = c('nzv', 'knnImpute'), k = 5)

# Create processed predictors
trnXSDI <- predict(trnXSDpp, newdata = trnXSD)
```

```{r runModel1Sev}
trC <- trainControl(method = 'none')
set.seed(878)
m1SevFit <- train(x = trnXSDI, y = trnSev, trControl = trC,
                  method = 'glmStepAIC', family = gaussian, direction = 'both',
                  trace = 0)
kable(summary(m1SevFit)$coefficients, digits = 3L,
      caption = "Model 1 Severity Regression Output")
```

@@

# SELECT MODELS
## Model Selection Criteria
For that *classification* algorithm, we will look at accuracy, AUC, and F1, and
select the model which performs best two out of three in those metrics. For the
*regression* algorithm we will look at RMSE and MAE (since the values must
be positive) and select the model which performs best on both, Ties will be
broken by adjusted \(R^2\).

```{r setUpCompTables}
freqCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                            ACC = double(3),
                            F1 = double(3),
                            AUC = double(3))
sevCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                           RMSE = double(3),
                           MAE = double(3),
                           adjR2 = double(3))
```

## Model 1
### Frequency Model

```{r validateM1F}
# Model 1 Frequency:: process validation data similar to training data
tstFreq <- factor(tstX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
tstDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_AMT,
                        fullRank = TRUE, data = tstX)
tstXFD <- predict(tstDumFreq, newdata = tstX)
tstXFDpp <- preProcess(tstXFD, method = c('nzv', 'knnImpute'))
tstXFDI <- predict(tstXFDpp, newdata = tstXFD)

# Model 1 Frequency: predict on validation data
m1Fpred <- predict(m1FreqFit, newdata = tstXFDI)
m1FpredProb <- predict(m1FreqFit, newdata = tstXFDI, type = 'prob')

# Model 1 Frequency: Calculate accuracy statistics
m1FCM <- confusionMatrix(m1Fpred, tstFreq, mode = 'everything')
freqCompTable[1, 2] <- m1FACC <- sum(diag(m1FCM$table)) / sum(m1FCM$table)
freqCompTable[1, 3] <- m1FF1 <- m1FCM$byClass[7]
freqCompTable[1, 4] <- m1FAUC <- auc(response = tstFreq,
                                     predictor = m1FpredProb[, 1])
```

Frequency Model 1  returns an accuracy of `r round(m1FACC, 4)`, an F1 score of
`r round(m1FF1, 4)`, and an AUC of `r round(m1FAUC, 4)`.

### Severity Model

```{r validateM1S}
# Model 1 Severity: process validation data similar to training data
tstSev <- tstX[TARGET_AMT > 0, TARGET_AMT]
nobstst <- length(tstSev)
np <- summary(m1SevFit$finalModel)$df[3]
tstDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                          TARGET_FLAG,
                        fullRank = TRUE, data = tstX[TARGET_AMT > 0])
tstXSD <- predict(tstDumSev, newdata = tstX[TARGET_AMT > 0])
tstXSDpp <- preProcess(tstXSD, method = c('nzv', 'knnImpute'), k = 5)
tstXSDI <- predict(tstXSDpp, newdata = tstXSD)

# Model 1 Severity: predict on validation data
m1Spred <- predict(m1SevFit, newdata = tstXSDI)

# Model 1 Severity: Calculate accuracy statistics
m1Ssum <- postResample(m1Spred, tstSev)
sevCompTable[1, 2] <- m1SRMSE <- m1Ssum[1]
sevCompTable[1, 3] <- m1SMAE <- m1Ssum[3]
m1SAR2 <- 1 - (1 - m1Ssum[2]) * (nobstst - 1) / (nobstst - np - 1)
sevCompTable[1, 4] <- m1SAR2
```

Severity Model 1  returns an RMSE of `r round(m1SRMSE, 1)`, an MAE of
`r round(m1SMAE, 1)`, and an adjusted \(R^2\) of `r round(m1SAR2, 4)`.

## Model 2
### Frequency Model
### Severity Model

## Model 3
### Frequency Model
### Severity Model

## Model Selection
### Frequency Model
```{r freqComp}
kable(freqCompTable, digits = 4L, caption = "Frequency Model Comparison")
```

### Severity Model
```{r sevComp}
kable(sevCompTable, digits = 4L, caption = "Severity Model Comparison")
```

@@

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r zeroVal}
```
```{r statsN}
```
```{r graphsN}
```
```{r graphsF}
```
```{r corrgram}
```
```{r dataClean}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r model1DummyFreq}
```
```{r model1RunFreq}
```
```{r model1DummySev}
```
```{r model1RunSev}
```
```{r setUpCompTables}
```
```{r validateM1F}
```
```{r validateM1S}
```
```{r freqComp}
```
```{r sevComp}
```
